---
title: "Examples of Hierarchical Latent Class Models for Mortality Surveillance Using Partially Verified Verbal Autopsies"
author: "Yu (Zoey) Zhu"
date: "2025-01-13"
output: pdf_document
geometry: margin=1in
---

```{r}
# Import packages
library(Rcpp)
library(MCMCpack)
library(ggplot2)
library(MetBrewer)
library(plyr)
library(MASS)
library(mnormt)
library(BayesLogit)
library(gsignal)
library(gtools)
library(tidyr)
library(caret)
```
# Simulation without aggregation example

In this report we show one simple example of simulation study without aggregations. The processes are (1) loading the model fitting functions and simulated data generating function. (2) setting up the model parameters and verification mechanism argument. (3) generating true prevalence and simulate the data. (4) fitting the models. (5) visualize the results, i.e, posterior mean and CI for the models comparing with the simulated true prevalence.
The simulation study with aggregations to the time and population levels examples are attached in `analysis_no_aggregate_sim_data.R`.


Import functions:

```{r}
## Rcpp dependencies
sourceCpp("../model/Rcpp/fit_model_BL.cpp")
sourceCpp("../model/Rcpp/fit_model_with_structured_prior.cpp")
## Models
source("../model/fit_model_BL.R")
source("../model/fit_model_with_structured_prior.R")
## Data generation function
source("../simulation/sim_with_sex_time_age.R")
```

Set up the model parameters and verification mechanism argument:

```{r}
# Set up 
n_Sex = 2
n_Age = 8
n_Time = 10

n_subsize = 100 # sample size of each sub-population
q = 10
K = 10

DEPEND_ON_Y = FALSE # Case (i)

```


Generate pi_true as prevalence: 

```{r}
p_S = c(0.2, 0.3)

x <- 1:n_Time
p_T <- (-6 + 10*x - x^2) / n_Time / 2

x <- 1:n_Age
p_A = (-6 + 10*x - x^2) / n_Age / 2

Nsta = n_Sex * n_Time * n_Age
pi_true = rep(NA, Nsta)
S_ind = c(rep(0, n_Time * n_Age), rep(1, n_Time*n_Age))
T_ind =  rep(rep(1:n_Time, each = n_Age), n_Sex)
A_ind =  rep(rep(1:n_Age,  n_Time), n_Sex)
for(i in 1:Nsta){
  pi_true[i] = round(expit(-1 + p_S[S_ind[i]+1] + p_T[T_ind[i]] + p_A[A_ind[i]]), 2) 
}
```

Fit the models:

```{r}
# generate simulated dataset
sim_data = sim_with_sex_time_age(pi_true, n_subsize = n_subsize, q = q, n_Sex = n_Sex, n_Time = n_Time, n_Age = n_Age, K = K, DEPEND_ON_Y = DEPEND_ON_Y)
  
E = 2000
BURN_IN = 1000
# fit models
fitted_model_BL = fit_model_BL(sim_data, E = E, BURN_IN = BURN_IN)
fitted_model_Fixed = fit_model_with_structured_prior(sim_data, K = K, E = E, BURN_IN = BURN_IN, STRUCTURED_PRIOR = "Fixed")
fitted_model_Indep = fit_model_with_structured_prior(sim_data, K = K, E = E, BURN_IN = BURN_IN, STRUCTURED_PRIOR = "Indep")
fitted_model_RW = fit_model_with_structured_prior(sim_data, K = K, E = E, BURN_IN = BURN_IN, STRUCTURED_PRIOR = "RW")
  
```

Visualize the posterior mean and CI of prevalence for the models:

```{r}
est0 <- data.frame(sex = c(rep("F", n_Time * n_Age), rep("M", n_Time * n_Age)),
                     time = rep(rep(1:n_Time, each = n_Age), 2), 
                     age = rep(rep(1:n_Age, n_Time), 2), 
                     prev = fitted_model_BL$bias+sim_data$param$pi, 
                     lower = apply(fitted_model_BL$poster_sample$sample_pis,2, function(x) quantile(x, 0.025)),
                     upper = apply(fitted_model_BL$poster_sample$sample_pis,2, function(x) quantile(x, 0.975)),
                     type = "BL")
est1 <- data.frame(sex = c(rep("F", n_Time * n_Age), rep("M", n_Time * n_Age)),
                     time = rep(rep(1:n_Time, each = n_Age), 2), 
                     age = rep(rep(1:n_Age, n_Time), 2), 
                     prev = fitted_model_Fixed$bias+sim_data$param$pi, 
                     lower = apply(fitted_model_Fixed$poster_sample$sample_pis,2, function(x) quantile(x, 0.025)),
                     upper = apply(fitted_model_Fixed$poster_sample$sample_pis,2, function(x) quantile(x, 0.975)),
                     type = "Fixed")
est2 <-  data.frame(sex = c(rep("F", n_Time * n_Age), rep("M", n_Time * n_Age)),
                      time = rep(rep(1:n_Time, each = n_Age), 2), 
                      age = rep(rep(1:n_Age, n_Time), 2), 
                      prev = fitted_model_Indep$bias+sim_data$param$pi, 
                      lower = apply(fitted_model_Indep$poster_sample$sample_pis,2, function(x) quantile(x, 0.025)),
                      upper = apply(fitted_model_Indep$poster_sample$sample_pis,2, function(x) quantile(x, 0.975)),
                      type = "Indep")
est3 <-  data.frame(sex = c(rep("F", n_Time * n_Age), rep("M", n_Time * n_Age)),
                      time = rep(rep(1:n_Time, each = n_Age), 2), 
                      age = rep(rep(1:n_Age, n_Time), 2), 
                      prev = fitted_model_RW$bias+sim_data$param$pi, 
                      lower = apply(fitted_model_RW$poster_sample$sample_pis,2, function(x) quantile(x, 0.025)),
                      upper = apply(fitted_model_RW$poster_sample$sample_pis,2, function(x) quantile(x, 0.975)),
                      type = "RW")


param <- data.frame(sex = c(rep("F", n_Time * n_Age), rep("M", n_Time * n_Age)),
                      time = rep(rep(1:n_Time, each = n_Age), 2), 
                      age = rep(rep(1:n_Age, n_Time), 2), 
                      prev = sim_data$param$pi, 
                      lower = NA,
                      upper = NA)

emp.obs <- aggregate(sim_data$data$Y.t ~sim_data$data$A + sim_data$data$Ti + sim_data$data$S  , FUN = mean, na.rm = TRUE)

observed <- data.frame(sex = c(rep("F", n_Time * n_Age), rep("M", n_Time * n_Age)),
                         time = rep(rep(1:n_Time, each = n_Age), 2), 
                         age = rep(rep(1:n_Age, n_Time), 2), 
                         prev = emp.obs$`sim_data$data$Y.t`, 
                         lower = NA,
                         upper = NA)


out <- rbind(est0, est1, est2, est3)

  time_index = month.abb[1:10]
  out$time <- time_index[out$time]
  out$time <- factor(out$time, levels = time_index)
  param$time <- time_index[param$time]
  param$time <- factor(param$time, levels = time_index)

  observed$time <- factor(observed$time, levels = 1:n_Time)
  out$age <- factor(out$age, levels = 1:n_Age)
  param$age <- factor(param$age, levels = 1:n_Age)
  observed$age <- factor(observed$age, levels = 1:n_Age)
  out$sex <- factor(out$sex, levels = c("F", "M"))
  param$sex <- factor(param$sex, levels = c("F", "M"))
  observed$sex <- factor(observed$sex,  levels = c("F", "M"))
  out$type <- factor(out$type, levels = c("BL", "Fixed", "Indep","RW"))


  out$type <- revalue(out$type, c("BL" = "Unstructured", 
                                  "Fixed" = "Fixed Effect", 
                                  "Indep" = "Indep RE", 
                                  "RW" = "RW1 RE"))
  out$sex <- revalue(out$sex, c("M" = "Male", 
                                "F" = "Female"))
  param$sex <- revalue(param$sex, c("M" = "Male", 
                                "F" = "Female"))

  cols <- met.brewer("Juarez", n = 6, type = "discrete")[-4]


g_posterior <- ggplot(subset(out, type %in% c("Unstructured", "RW1 RE")), aes(x = age, y = prev))  + 
    geom_ribbon(aes(ymin = lower, ymax = upper, fill = type, group = interaction(time, type)), alpha = 0.4) + 
    geom_line(aes(color = type, group = interaction(time, type)), linewidth = 1.05) + 
    geom_point(data = param, aes(x = age, y = prev), col = "black", size = 1.2) + 
    facet_grid(sex ~ time) +
    scale_color_manual("Model", values = c("#d95f02", "#1b9e77"))+ 
    scale_fill_manual("Model", values = c("#d95f02", "#1b9e77"))+ 
    theme_bw() +
    xlab("Age Group") + ylab("Prevalence") +
    theme(axis.text.x = element_text(size = 10), # Make X axis labels (age) bold
          strip.text = element_text(face = "bold",size = 10),  # Make facet labels (time) bold
          legend.text = element_text(size = 10))+ 
    theme(legend.position = "bottom", strip.text = element_text(size = 12))

print(g_posterior)
```


